---
title: "Practical 1 Answers"
subtitle: "Working with Data: Tools for Collaborating on a Project"
author: "Your Name Here"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: false
    link-external-newwindow: true
    link-external-icon: true
editor: source
---

<!--
Before you start:

- Make sure this file lives in the root of your project repository, alongside
  a `data/` folder containing pixar-films.csv and pixar-films-response.csv
  (see the practical instructions for the download links).
- See the README.md in this repository for package installation instructions
  and Git/GitHub tips.
- Render this document with the "Render" button in RStudio, or by running
  quarto::quarto_render("practical1-answers.qmd") in the R console.
-->

## Getting started

```{r}
#| label: setup
#| message: false
#| warning: false

# Load the packages you need for this practical.
# See the README for installation instructions if any of these are missing.
library(readr)
library(dplyr)
library(naniar)
library(ggplot2)
library(tibble)
```

## Reading in data

### Question 1

Download the `.csv` files and use `readr::read_csv()` to read each dataset into R.

```{r}
#| label: read-data

# pixar_films <- read_csv("data/pixar-films.csv")
# public_response <- read_csv("data/pixar-films-response.csv")
```

### Question 2

Inspect each dataset. What data is missing? Is the missingness coded uniformly? Is it structural missingness, or can it be corrected?

```{r}
#| label: inspect-data

```





## Working with missingness

### Question 3

Use functions from `naniar` to summarise the missingness for each dataset. Does it capture all the missing data?

```{r}
#| label: missingness-summary

```





### Question 4

Clean up the data. You may find row manipulations from the `dplyr` package useful (e.g. `dplyr::rows_update()`).

```{r}
#| label: clean-data

```

### Question 5

Now that they are clean, merge the two datasets using one of the `dplyr::*_join()` functions. Does it matter which join function you use? Why/why not?

```{r}
#| label: merge-data

```





## Bias

### Question 6

What bias is present in the dataset? Can it be corrected?





### Question 7

If you identified bias that can be corrected, add it (e.g. using `dplyr::rows_insert()`, similar to `dplyr::rows_update()`).

```{r}
#| label: correct-bias

```

### Question 8

Collect some additional data for each movie from the internet (e.g. Wikipedia), and add it to the merged dataset. Create a new dataset (either as a `.csv` or directly in your script) and join it in with `dplyr::left_join()`.

```{r}
#| label: additional-data

```

## Visualisation

### Question 9

Create plots of the data that visualise two different aspects of the data.

```{r}
#| label: plot-1

```

```{r}
#| label: plot-2

```

## Finishing up

### Question 10

Is your code transparent and understandable? Make sure the chunks above are commented to explain the purpose and techniques used, and that this document includes a short description of what it does (see the top of this file).

### Question 11

Can your code run as a reproducible standalone script (i.e. no user input)? Check that it runs top-to-bottom in a fresh R session (`Restart R` then `Render`).

### Question 12

Ensure you save a copy of the raw data, new data, and script together, and write your tidy data to file with `readr::write_csv()`.

```{r}
#| label: save-output
#| eval: false

# write_csv(merged_data, "data/pixar-films-clean.csv")
```
