---
title: "Practical 1 Answers"
subtitle: "Working with Data: Tools for Collaborating on a Project"
author: "Tia Light"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: false
    link-external-newwindow: true
    link-external-icon: true
editor: source
---

<!--
Before you start:

- Make sure this file lives in the root of your project repository, alongside
  a `data/` folder containing pixar-films.csv and pixar-films-response.csv
  (see the practical instructions for the download links).
- See the README.md in this repository for package installation instructions
  and Git/GitHub tips.
- Render this document with the "Render" button in RStudio, or by running
  quarto::quarto_render("practical1-answers.qmd") in the R console.
-->

## Getting started

```{r}
#| label: setup
#| message: false
#| warning: false

# Load the packages you need for this practical.
# See the README for installation instructions if any of these are missing.
library(readr)
library(dplyr)
library(naniar)
library(ggplot2)
library(tibble)
```

## Reading in data

### Question 1

Download the `.csv` files and use `readr::read_csv()` to read each dataset into R.

```{r}
#| label: read-data

pixar_films <- read_csv("data/pixar-films.csv")
public_response <- read_csv("data/pixar-films-response.csv")
```

### Question 2

Inspect each dataset. What data is missing? Is the missingness coded uniformly? Is it structural missingness, or can it be corrected?

```{r}
#| label: inspect-data


pixar_films
pixar_films |> summarise(across(everything(), ~ sum(is.na(.))))
pixar_films |> summarise(across(c(number,film_rating), ~ sum(. %in% c("Not Rated", "N/A"))))

# Correct N/A coding
pixar_films <- pixar_films |> 
  mutate(film_rating = na_if(film_rating, "N/A")) |> 
  mutate(film_rating = na_if(film_rating, "Not Rated"))

# outliers in runtime
ggplot(pixar_films) + geom_histogram(aes(x = run_time))

# replace outliers with NAs
pixar_films <- pixar_films |> mutate(run_time = replace_when(run_time, run_time > 145 ~ NA))

# inspect all NAs
pixar_films |> filter(if_any(everything(), is.na))

public_response
# correctly coded missing data in all columns except film
public_response |> summarise(across(everything(), ~ sum(is.na(.))))

```





## Working with missingness

### Question 3

Use functions from `naniar` to summarise the missingness for each dataset. Does it capture all the missing data?

```{r}
#| label: missingness-summary

naniar::miss_var_summary(pixar_films)
naniar::vis_miss(pixar_films)

naniar::miss_var_summary(public_response)
naniar::vis_miss(public_response)


```





### Question 4

Clean up the data. You may find row manipulations from the `dplyr` package useful (e.g. `dplyr::rows_update()`).

```{r}
#| label: clean-data

# The unnamed film is Elemental, identifiable by its release date.
# Use dplyr::rows_update() to correct the known errors/fill in known values,
# matching by release_date since film itself is one of the missing values.
pixar_films <- pixar_films |>
  rows_update(
    tibble(
      release_date = as.Date(c(
        "2021-06-18", # Luca
        "2022-03-11", # Turning Red
        "2022-06-17", # Lightyear
        "2023-06-16"  # Elemental
      )),
      film = c("Luca", "Turning Red", "Lightyear", "Elemental"),
      run_time = c(95, 100, 105, 103),
      film_rating = c("PG", "PG", "PG", "PG")
    ),
    by = "release_date"
  )

pixar_films

```

### Question 5

Now that they are clean, merge the two datasets using one of the `dplyr::*_join()` functions. Does it matter which join function you use? Why/why not?

```{r}
#| label: merge-data


# left_join() keeps every film in pixar_films (the "complete" list of 27
# films) and fills in NA for any film with no matching row in
# public_response.
pixar_data <- left_join(pixar_films, public_response, by = "film")

pixar_data

naniar::vis_miss(pixar_data)


```





## Bias

### Question 6

What bias is present in the dataset? Can it be corrected?

The clearest bias is a temporal bias: the three most recently released films (Turning Red, Lightyear, Elemental) are absent from public_response, so any summary of critical/audience reception computed from the data would silently exclude these outputs. This is at least partly correctable, since real critic/audience scores exist for these films.

More recent Pixar films (e.g. Toy Story 5) are also missing from the dataset!

### Question 7

If you identified bias that can be corrected, add it (e.g. using `dplyr::rows_insert()`, similar to `dplyr::rows_update()`).

```{r}
#| label: correct-bias

# Real critic/audience scores for the three missing films, collected from
# Rotten Tomatoes and Metacritic (accessed 2026-08).
missing_response <- tribble(
  ~film,          ~rotten_tomatoes, ~metacritic, ~cinema_score, ~critics_choice,
  "Turning Red",  95,               83,          NA,            NA,
  "Lightyear",    74,               60,          "A",           NA,
  "Elemental",    75,               59,          "A",           NA
)

public_response <- public_response |>
  rows_insert(missing_response, by = "film")

# Re-do the merge now that public_response is more complete.
pixar_data <- left_join(pixar_films, public_response, by = "film")

pixar_data
```

### Question 8

Collect some additional data for each movie from the internet (e.g. Wikipedia), and add it to the merged dataset. Create a new dataset (either as a `.csv` or directly in your script) and join it in with `dplyr::left_join()`.

```{r}
#| label: additional-data

# Directors for each film, collected from Wikipedia.
directors <- tribble(
  ~film,                   ~director,
  "Toy Story",             "John Lasseter",
  "A Bug's Life",          "John Lasseter",
  "Toy Story 2",           "John Lasseter",
  "Monsters, Inc.",        "Pete Docter",
  "Finding Nemo",          "Andrew Stanton",
  "The Incredibles",       "Brad Bird",
  "Cars",                  "John Lasseter",
  "Ratatouille",           "Brad Bird",
  "WALL-E",                "Andrew Stanton",
  "Up",                    "Pete Docter",
  "Toy Story 3",           "Lee Unkrich",
  "Cars 2",                "John Lasseter",
  "Brave",                 "Mark Andrews",
  "Monsters University",   "Dan Scanlon",
  "Inside Out",            "Pete Docter",
  "The Good Dinosaur",     "Peter Sohn",
  "Finding Dory",          "Andrew Stanton",
  "Cars 3",                "Brian Fee",
  "Coco",                  "Lee Unkrich",
  "Incredibles 2",         "Brad Bird",
  "Toy Story 4",           "Josh Cooley",
  "Onward",                "Dan Scanlon",
  "Soul",                  "Pete Docter",
  "Luca",                  "Enrico Casarosa",
  "Turning Red",           "Domee Shi",
  "Lightyear",             "Angus MacLane",
  "Elemental",             "Peter Sohn"
)

pixar_data <- pixar_data |>
  left_join(directors, by = "film")

pixar_data

```

## Visualisation

### Question 9

Create plots of the data that visualise two different aspects of the data.

```{r}
#| label: plot-1

# Aspect 1: has critical reception (Rotten Tomatoes score) changed over time?
ggplot(pixar_data, aes(x = release_date, y = rotten_tomatoes)) +
  geom_point() +
  scale_x_date("Release date") +
  scale_y_continuous("Rotten Tomatoes score (%)") +
  ggtitle("Pixar films' critical reception over time") +
  theme_bw()


```

```{r}
#| label: plot-2

# Aspect 2: how does runtime relate to age rating?
ggplot(pixar_data, aes(x = film_rating, y = run_time)) +
  geom_boxplot() +
  geom_jitter(width = 0.1, alpha = 0.6) +
  scale_x_discrete("Film rating") +
  scale_y_continuous("Run time (mins)") +
  ggtitle("Runtime by age rating") +
  theme_bw()

```

## Finishing up

### Question 10

Is your code transparent and understandable? Make sure the chunks above are commented to explain the purpose and techniques used, and that this document includes a short description of what it does (see the top of this file).

### Question 11

Can your code run as a reproducible standalone script (i.e. no user input)? Check that it runs top-to-bottom in a fresh R session (`Restart R` then `Render`).

### Question 12

Ensure you save a copy of the raw data, new data, and script together, and write your tidy data to file with `readr::write_csv()`.

```{r}
#| label: save-output
#| eval: false

# write_csv(merged_data, "data/pixar-films-clean.csv")
```
